---
title: "Assignment1"
output:
  html_document:
    df_print: paged
  word_document: default
date: "2022-11-11"
---

```{r}

rm(list=ls())  

library(psych) #fa()
library(lavaan) #cfa()
library(MASS)
library(paran) #procedure Horn

#function composite reliability
compositerel<-function(x){
  A<-(sum(x))^2
  B<-sum(1-x^2)
  return(A/(A+B))
}


#load data
load("/Users/yavuzhanyavuz/Desktop/Leuven/Multivariate/cosmetics.Rdata")

#compute centered data
ccosmetics<-cosmetics
ccosmetics[,1:18]<-scale(cosmetics[,1:18],center=TRUE,scale=FALSE)

```

## a. Use CFA to construct a measurement model for the Attitude items

```{r}
#step1:confirmatory factor analysis model with 3 correlated latent variables

#measurement model latent variables
cfa1<-' Att_organic=~1*Attitude_organic1+Attitude_organic2+Attitude_organic3         
        Att_packaging=~1*Attitude_packaging1+Attitude_packaging2+Attitude_packaging3
        Att_crueltyfree=~1*Attitude_crueltyfree1+Attitude_crueltyfree2+Attitude_crueltyfree3
        '

#fit model on covariance matrix
fitcfa1<-cfa(cfa1,ccosmetics[,1:9])

#print fitmeasures
fitmeasures(fitcfa1,c("chisq","df","pvalue","cfi","tli","rmsea","srmr"))

#summary of results
summary(fitcfa1,fit.measures=TRUE)

#print standardized solution
standardizedSolution(fitcfa1)


#reliability factor scores
d<-standardizedSolution(fitcfa1)
#composite reliability attitude_organic
compositerel(d[1:3,4])
#composite reliability attitude_packaging
compositerel(d[4:6,4])
#composite reliability attitude_crueltyfree
compositerel(d[7:9,4])

#overview table composite reliability
factorscore<-c("attitude_organic","attitude_packaging","attitude_crueltyfree")
reliability<-round(c(compositerel(d[1:3,4]),compositerel(d[4:6,4]),compositerel(d[7:9,4])),3)
data.frame(factorscore,reliability)

## step2: correlated error terms for variables measured using the same method #!
cfa2<-' Att_organic=~1*Attitude_organic1+Attitude_organic2+Attitude_organic3         
        Att_packaging=~1*Attitude_packaging1+Attitude_packaging2+Attitude_packaging3
        Att_crueltyfree=~1*Attitude_crueltyfree1+Attitude_crueltyfree2+Attitude_crueltyfree3
        
        Att_organic ~~1*Att_organic
        Att_packaging ~~ 1*Att_packaging
        Att_crueltyfree ~~1*Att_crueltyfree
        
        Att_organic ~~ Att_packaging
        Att_packaging ~~ Att_crueltyfree
        Att_organic ~~ Att_crueltyfree
        
        Attitude_organic1~~a*Attitude_packaging1
        Attitude_organic1~~a*Attitude_crueltyfree1
        Attitude_packaging1~~a*Attitude_crueltyfree1
        
        Attitude_organic2~~b*Attitude_packaging2
        Attitude_organic2~~b*Attitude_crueltyfree2
        Attitude_packaging2~~b*Attitude_crueltyfree2
        
        Attitude_organic3~~c*Attitude_packaging3
        Attitude_organic3~~c*Attitude_crueltyfree3
        Attitude_packaging3~~c*Attitude_crueltyfree3
        '
#fit model on covariance matrix
fitcfa2<-cfa(cfa2,ccosmetics)

#print fitmeasures
fitmeasures(fitcfa2,c("chisq","df","pvalue","cfi","tli","rmsea","srmr"))

#summary of results
summary(fitcfa2,fit.measures=TRUE)


lavInspect(fitcfa2, "cor.lv")
#print standardized solution
standardizedSolution(fitcfa2)

```

## b. Use CFA to construct a measurement model for the Behavior-Intention items

After loading the data, we compute centered variables. We fit a CFA model with 3 correlated latent variables for the and print fit measures and the standardized solution.

```{r}
#step1:confirmatory factor analysis model with 3 correlated latent variables
cfa3<-' BI_organic=~1*BI_organic1+BI_organic2+BI_organic3         
        BI_packaging=~1*BI_packaging1+BI_packaging2+BI_packaging3
        BI_crueltyfree=~1*BI_crueltyfree1+BI_crueltyfree2+BI_crueltyfree3
        '

#fit model on covariance matrix
fitcfa3<-cfa(cfa3,ccosmetics)

#print fitmeasures
fitmeasures(fitcfa3,c("chisq","df","pvalue","cfi","tli","rmsea","srmr"))

#summary of results
summary(fitcfa3,fit.measures=TRUE)

#print standardized solution
standardizedSolution(fitcfa3)
```

The **fit measures** indicate that the model is rejected by an absolute goodness of fit test, i.e. the fit of the model is significantly lower than for a perfectly fitting model (chi-square=147.8, df=24, p\<.001). Descriptive fit measures indicate that the model does not fit the covariance matrix well: CFI (.914) and TLI (.871) both below the cutoff of .95 which indicates good fit. In addition, RMSEA (.185) and SRMR (.033) so that the cutoff value for good model fit(cutoff\<.08) is only satisfied for SRMR. Given these results, further modifications to the model are needed.

**Factor correlations**:From the covariances matrix, we can see the factors have positive significant correlations.

As can be seen in the standardized solution, all variables have significant and positive standardized loadings that exceed 0.7. Hence, the variables have sufficient reliability so that **convergent validity** is satisfied for the measurement model. Furthermore, **divergent validity** is also satisfied as all latent variables have moderate correlations that are significantly smaller than 1.

Finally, the composite reliability of all the factor scores is good as it exceeds .80.

```{r}
#reliability factor scores
d<-standardizedSolution(fitcfa3)
#composite reliability BI_organic
compositerel(d[1:3,4])
#composite reliability BI_packaging
compositerel(d[4:6,4])
#composite reliability BI_crueltyfree
compositerel(d[7:9,4])


#overview table composite reliability
factorscore<-c("BI_organic","BI_packaging","BI_crueltyfree")
reliability<-round(c(compositerel(d[1:3,4]),compositerel(d[4:6,4]),compositerel(d[7:9,4])),3)
data.frame(factorscore,reliability)
```

Next we extend the CFA model by including correlated error terms, which we imposed the constraint of equal residual correlations, for all pairs of items that focus on the same aspect.

```{r}
## step2: correlated error terms for variables measured using the same method
cfa4<-' BI_organic=~1*BI_organic1+BI_organic2+BI_organic3         
        BI_packaging=~1*BI_packaging1+BI_packaging2+BI_packaging3
        BI_crueltyfree=~1*BI_crueltyfree1+BI_crueltyfree2+BI_crueltyfree3
        
        BI_organic ~~1*BI_organic
        BI_packaging ~~ 1*BI_packaging
        BI_crueltyfree ~~1*BI_crueltyfree
        
        BI_organic ~~ BI_packaging
        BI_packaging ~~ BI_crueltyfree
        BI_organic ~~ BI_crueltyfree
        
        BI_organic1~~a*BI_packaging1
        BI_organic1~~a*BI_crueltyfree1
        BI_packaging1~~a*BI_crueltyfree1
        
        BI_organic2~~b*BI_packaging2
        BI_organic2~~b*BI_crueltyfree2
        BI_packaging2~~b*BI_crueltyfree2
        
        BI_organic3~~c*BI_packaging3
        BI_organic3~~c*BI_crueltyfree3
        BI_packaging3~~c*BI_crueltyfree3
        '
#fit model on covariance matrix
fitcfa4<-cfa(cfa4,ccosmetics[,10:18])

#print fitmeasures
fitmeasures(fitcfa4,c("chisq","df","pvalue","cfi","tli","rmsea","srmr"))

#summary of results
summary(fitcfa4,fit.measures=TRUE)

#print standardized solution
standardizedSolution(fitcfa4)
```

The **fit measures** of the extended CFA model indicate that the model fits the data well(chi-square=31.251, df=24, p=.147) and has excellent descriptive goodness of fit.CFI=.995 and TLI=.992 both exceed the cutoff of .95 which indicates good fit, RMSEA (.045) and SRMR (.074) also indicate a good fit as they are well below the cutoff of .08. Given the results, the fit measures of the CFA models fitted in the second step is much better than the first step.

The results of the correlated error model indicate that items focus on the same aspect have significant positive **residual correlations**. i.e. for "making an effort to buy", the residual correlation is around .33; for "recommending", the residual correlation is around 0.51; for "checking the sustainable label", the residual correlation is between .22-.32.

## c.Build a structural equation model to evaluate the impact of attitude on behavior intention

#We use the **sem()** function to fit the structural equation model on the covariance matrix, and print fit measures and model output (including the standardized solution).

```{r}
##first step
sem1 <-'# measurement model
        Att_organic=~Attitude_organic1+Attitude_organic2+Attitude_organic3         
        Att_packaging=~Attitude_packaging1+Attitude_packaging2+Attitude_packaging3
        Att_crueltyfree=~Attitude_crueltyfree1+Attitude_crueltyfree2+Attitude_crueltyfree3
        BI_organic=~1*BI_organic1+BI_organic2+BI_organic3         
        BI_packaging=~1*BI_packaging1+BI_packaging2+BI_packaging3
        BI_crueltyfree=~1*BI_crueltyfree1+BI_crueltyfree2+BI_crueltyfree3
        
        # structural model(regressions)
        BI_organic~Att_organic
        BI_packaging~Att_packaging
        BI_crueltyfree~Att_crueltyfree
        
        #variance latent variables
        Att_organic~~Att_organic
        Att_packaging~~Att_packaging
        Att_crueltyfree~~Att_crueltyfree
        BI_organic~~1*BI_organic
        BI_packaging~~1*BI_packaging
        BI_crueltyfree~~1*BI_crueltyfree'

fitsem1 <- sem(sem1, data =ccosmetics)
summary(fitsem1, fit.measure = TRUE)
standardizedSolution(fitsem1)
fitmeasures(fitsem1,c("chisq","df","pvalue","cfi","tli","rmsea","srmr"))
```

```{r}
##second step
sem2 <-'# measurement model
        Att_organic=~Attitude_organic1+Attitude_organic2+Attitude_organic3         
        Att_packaging=~Attitude_packaging1+Attitude_packaging2+Attitude_packaging3
        Att_crueltyfree=~Attitude_crueltyfree1+Attitude_crueltyfree2+Attitude_crueltyfree3
        BI_organic=~1*BI_organic1+BI_organic2+BI_organic3         
        BI_packaging=~1*BI_packaging1+BI_packaging2+BI_packaging3
        BI_crueltyfree=~1*BI_crueltyfree1+BI_crueltyfree2+BI_crueltyfree3
        
        # structural model(regressions)
        BI_organic~a*Att_organic
        BI_packaging~a*Att_packaging
        BI_crueltyfree~a*Att_crueltyfree
        
        #variance latent variables
        Att_organic~~Att_organic
        Att_packaging~~Att_packaging
        Att_crueltyfree~~Att_crueltyfree
        BI_organic~~1*BI_organic
        BI_packaging~~1*BI_packaging
        BI_crueltyfree~~1*BI_crueltyfree'


fitsem2 <- sem(sem2, data =ccosmetics)
summary(fitsem2, fit.measure = TRUE)
standardizedSolution(fitsem2)
fitmeasures(fitsem2,c("chisq","df","pvalue","cfi","tli","rmsea","srmr"))
```

The results indicates that the step2 model is still rejected by a good ness of fit test(chi-square=389.9, df=131, p\<.001). And for the descriptive fit, the model does not fit the covariance matrix well(CFI=.897, TLI=.880, RMSEA=.115, SRMR=.194).

From the result, we can see that there is a significant positive effect of importance of attitude on behavior intention. With a standard deviation increase in each of the attention score, the score on behavior intention factor will each increase by 11.41 standard deviation.

```{r}
#model fit comparison
anova(fitsem1, fitsem2)
```

Since we accept the null hypothesis using the chi-squared difference of fit test, this means that we do not need to impose the constraint that the 3 population regression coefficients of the structural model are equal.But according to the value of AIC and BIC, the second model is better than the first one.

From the standardized solution, we can find that all variables of behavior intention have significant and positive standardized loadings that exceed 0.7. But for attitude, only the reliabilities of att_organic1, att_crueltyfree1 and att_crueltyfree3 are exceed 0.7.

The factor score of Att_organic and Att_packaging have a composite reliability of .59 and .67, so they don't have acceptance reliability. But for Att_crueltyfree, BI_organic, BI_packaging and BI_crueltyfree, the composite reliabilities are 0.74, 0.83, 0.84 and 0.89, so they all have acceptance reliability.

Attitude items all have significant positive effect on behavior intention items(Att_organic on BI_organic, Att_packaging on BI_packaging and Att_Crueltyfree on BI_crueltyfree. The estimated correlations between each of them are 0.522, 0.530 and 0.573.

---
## Task 2 
### PART A
We load the data, standardize the variables, use the candisc() procedure to conduct
canonical correlation analysis and print a summary of the results and compute redundancies. 

```{r, results=FALSE,message=FALSE, warning=FALSE}
load("/Users/yavuzhanyavuz/Desktop/Leuven/Multivariate/benefits.Rdata") 

library(candisc)
#conduct canonical correlation analysis 

z_benefits = benefits ##first update, we need to normalize the datq
z_benefits[,2:14]<-scale(benefits[,2:14],center=TRUE,scale=TRUE)

cancor.out<-cancor(cbind(SL_pensioners,SL_unemployed,SL_old_gvntresp,SL_unemp_gvntresp) 
                   ~SB_strain_economy+SB_prevent_poverty+SB_equal_society  
                   +SB_taxes_business+SB_make_lazy+SB_caring_others  
                   +unemployed_notmotivated+SB_often_lessthanentitled
                   +SB_often_notentitled,data=z_benefits)
```
```{r}
summary(cancor.out) 

#computing redundancies from output 
R2tu<-cancor.out$cancor^2 
VAFYbyt<-apply(cancor.out$structure$Y.yscores^2,2,sum)/4 ##updated 5 to 4, now they match with redundancy function
redund<-R2tu*VAFYbyt 
round(cbind(R2tu,VAFYbyt,redund,total=cumsum(redund)),4)
```

The canonical correlation analysis extracts four pairs of canonical variates. Hypotheses tests
indicate that the fourth pair can be ignored as the canonical correlation is not significant, i.e., $H_0$: $\rho$($u_4$,$t_4$) = 0 cannot be rejected at the 5% level (p=.1735). 

The first canonical correlation equals 0.483. This means that the canonical variate $u_1$
accounts for 23.35% of the variance in the canonical variate $t_1$ . The second canonical
correlation equals 0.228. This means that the canonical variate $u_2$ accounts for 5.21% of the
variance in the canonical variate $t_2$.

As shown by the redundancies, the first three pairs of canonical variates account for 8.88% of
the variance in the Y variables. The first pair of canonical variates is most relevant at accounting for the variance in the Y variables by a score of 6.65%. The second pair of canonical variates contributes by 1.67% and the third accounts for only 0.5% in the explanation of the variance of Y. We decided to keep third canonical variate because it increases the total explanation of variance Y by 6% (0.0051/0.0832). In conclusion, three pairs are included in the interpretation of the relation but contributions decrease respectively with the pairs. 

Therefore, we will analyze the first pair of canonical variates. To interpret the first pair of canonical variates, we print the canonical loadings (=correlation between the canonical variates and the X and Y variables). In addition, we make a scatter plot of the first pair of canonical variates and indicate a different color for observations of each country. 

```{r}
#print canonical loadings 
round(cancor.out$structure$X.xscores,2)
round(cancor.out$structure$Y.yscores,2)
```

```{r, echo=FALSE}
#### plotting

colors <- c("#B4464B", # Red
            "#4682B4") # Steel blue
can1<-cbind(cancor.out$scores$X[,1],cancor.out$scores$Y[,1])
rownames(can1)<-as.character(z_benefits[,1])
plot(can1,xlab="u1",ylab="t1",xlim=c(-4,4),ylim=c(-4,4), lwd=0.6,col=colors[as.factor(rownames(can1))], pch=4)

legend("topleft",
       legend = c("Belgium", "Great Britain"),
       pch = 4,
       col = colors[factor(levels(as.factor(rownames(can1))))]) 
```
The canonical loadings indicate that Y variables are split on first two canonical variates based on question's topic, unemployed vs pensioners. $t_1$ has relatively higher correlation on two Y variables which are regarding unemployment of the people while $t_2$ has relatively higher correlation with questions regarding pensioners. Regarding the interpretation, high score on $t_1$ means that the respondent thinks standard living of unemployed people is poor and it is government's responsibility. On the other hand, $t_2$ is the inverse of $t_1$. **Low score** on $t_2$ means that the respondent believes standard living of pensioners is poor and it is government's responsibility. 

Similarly, $u_1$ is more correlated on questions about laziness and motivation of unemployed people. High score on $u_1$ means that respondent strongly disagrees with the idea that social benefits/services make people lazy and unemployment is their fault while also disagreeing that those people are unwilling to care for others when they receive benefits/services and obtain social benefits/services that are not entitled to them.

There is no clear distinction on first canonical variates between countries. Both countries have respondents for every point of view. But we can conclude that slight positive correlation between canonical variates align with respondent's perspective on both X and Y variables that have relatively higher correlation.

### PART B
```{r, message=FALSE, warning=FALSE}
#split data in two parts and standardize data 
train<-z_benefits[seq(2,nrow(z_benefits),by=2),] 
valid<-z_benefits[seq(1,nrow(z_benefits),by=2),] 
train[,2:14]<-scale(train[,2:14],center=TRUE,scale=TRUE) 
valid[,2:14]<-scale(valid[,2:14],center=TRUE,scale=TRUE) 

#conduct CCA on calibration data 
cancor.train<-cancor(cbind(SL_pensioners,SL_unemployed,SL_old_gvntresp,SL_unemp_gvntresp) 
                     ~SB_strain_economy+SB_prevent_poverty+SB_equal_society  
                     +SB_taxes_business+SB_make_lazy+SB_caring_others  
                     +unemployed_notmotivated+SB_often_lessthanentitled
                     +SB_often_notentitled,data=train)

#conduct CCA on validation data 
cancor.valid<-cancor(cbind(SL_pensioners,SL_unemployed,SL_old_gvntresp,SL_unemp_gvntresp) 
                     ~SB_strain_economy+SB_prevent_poverty+SB_equal_society  
                     +SB_taxes_business+SB_make_lazy+SB_caring_others  
                     +unemployed_notmotivated+SB_often_lessthanentitled
                     +SB_often_notentitled,data=valid) 

# canonical variates calibration set 
train.X1<-cancor.train$score$X 
train.Y1<-cancor.train$score$Y

# compute canonical variates using data of calibration set and 
#coefficients estimated on validation set 
train.Y2<-as.matrix(train[,2:5])%*%cancor.valid$coef$Y
train.X2<-as.matrix(train[,6:14])%*%cancor.valid$coef$X 
```

```{r}
#R(U*,T*) versus R(U,T) 
round(cor(train.X2,train.Y2)[1:4,1:4],3) 
round(cor(train.X1,train.Y1)[1:4,1:4],3) 
```
When diagonal elements are compared, it is seen thqt R(U$*$,T$*$) has slightly lower correlation values than R(U,T) for first two canonical variate pairs while the third one has a rather larger overestimation. The fourth pair may be skipped as it is also neglected in the model. Off-diagonal elements of R(U$*$,T$*$) are rather small and smaller in absolute values than diagonal elements. To sum up, overestimation due to maximization involved is not an issue.

```{r}
#R(T*,T*) and R(U*,U*) 
round(cor(train.Y2,train.Y2)[1:4,1:4],3) 
round(cor(train.X2,train.X2)[1:4,1:4],3) 
```
The off-diagonal elements of R(T$*$,T$*$) and R(U$*$,U$*$) are close to 0, which indicates that
canonical variates of Y variables and of X variables computed on calibration data but based
on the coefficients from validation data have as expected correlations that are close to 0. It can be concluded that they are independent.

## Part C

```{r}
#R(T,T*) and R(U,U*) 
round(cor(train.Y1,train.Y2)[1:4,1:4],3) 
round(cor(train.X1,train.X2)[1:4,1:4],3) 
```
The absolute value of the diagonal elements of R(T,T$*$) and R(U,U$*$) represent the reliabilities of the canonical variates for Y and X variables. The reliabilities of $t_1$, $t_2$, $t_3$, $t_4$ are .985, .989, .973 and .988, respectively. Also, the reliabilities of $u_1$, $u_2$, $u_3$, $u_4$ are as follows: .985, .893, .557 and .257. Subsequently, it can be concluded that first two pairs of canonical variates have acceptable reliability while remaining pairs have unacceptable reliability. Considering the redundancy analysis and canonical correlation hypothesis model, it is seen first two pairs are both important and reliable.

```
